
\section{Tests and benchmarks}
\label{tests_and_benchmarks}

There are 3 tests that can be performed:
\begin{itemize}
 \item The Dirac operator test (\verb|deo_doe_test|);
 \item The multishift inverter test (\verb|inverter_multishift_test|):
 \item The pure gauge MD test (main program, \verb|main|);
\end{itemize}

There is a script that will produce all the necessary slurm scripts, that is \\
\verb|scripts/prepare_slurm_benchmarks.sh|. \\ A way of calling it is the following:
\begin{verbatim}
cd build  # hopefully you already created this
../scripts/prepare_slurm_benchmarks.sh geom_defines.txt
\end{verbatim}
This script will create 
\begin{itemize}
 \item the slurm scripts for the 3 tests described
 \item the ``profiling-friendly'' versions of the 3 slurm scripts which perform order of 10 iterations.
 \item the 2 setting files needed for benchmarking
 \item other 2 setting files, ``profiling-friendly''
\end{itemize}

A brief description of the benchmarks:
\begin{itemize}
 \item {Dirac operator test}: \\
In the dirac operator test, first \verb|acc_Doe| is executed \verb|DeoDoeIterations| 
times, then the same for \verb|acc_Deo|, then the same for the whole Dirac operator 
which contains also a simple ({\sf saxpy} or {\sf daxpy}) linear combination step.
Notice that this benchmark can also be used to test the correctness of modifications to the
Dirac operator code.
\item {Multishift inverter test}: \\
The multishift inverter is called \verb|MultiShiftInverterRepetitions| times.
The parameters in the setting file for the benchmark are set so that all shifts 
will perform the same number of iterations (benchmark mode). The target residue is set very low so that the 
inverter will never converge in practice, and will stop at the maximum number of allowed iterations.
\item {Pure gauge test}: \\
In this case, the full program \verb|main| is used without fermions. The molecular dynamics is run 
in pure-gauge mode. Notice that stouting is performed, even if the fermion forces are exactly zero.
The important parameters are all in the {\sf MDParameters} group.
Notice that, with the diagnostics options, this test can be used to test the correcness
of modifications to the pure gauge part of Molecular dynamics.
\end{itemize}

The measured times in these benchmarks should be reproducible and ``gauge-configuration-invariant''.
Tests of the full program of course are reproducible but not ``gauge-configuration-invariant'' due
to the fact that, in normal conditions, the number of iterations in the multishift inverter varies 
depending on the gauge configuration. It is possible to actually change that behaviour, so that 
the multishift inverter will do a fixed number of iterations, but this possibility is at present not coded 
in the program.
